---
title: "effects"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters from Exp2 of the morality aesthetics project.

# load the libraries that we will be using #

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", 
         "tidybayes", "bayesplot", "future", "parallel", "flextable")

lapply(pkg, library, character.only = TRUE)
```

# adjust multicore settings #

```{r set-options}
options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()
```

# plot settings #

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face = "bold"), 
          title = element_text(size = 18, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# table settings #

```{r}
set_flextable_defaults(
  theme_fun = theme_vanilla,
  background.color = "white"
)
```

# read in previously saved brms models object files and/or fits #

If available, this saves time computing them each time 

```{r}
bf <- readRDS("exp2/models/bf.rds")
```

# read in the data #

```{r}
datadp <- read_csv("exp2/data/dprime.csv") %>%
  select(pid, groupd, targetd, attributed, dprime, ldprime) %>% 
  mutate(pid = factor(pid))
head(datadp)
str(datadp)
```

# section 1 #

## Look at the point and interval estimates for model bf (the full model) ##

using posterior_samples and tidybayes

```{r}
post <- as_draws_df(bf) %>%  #
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(bf, robust = TRUE)

# all fixed effects
post_qi_b <- post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_b)

# varying effects - SD
post_qi_sd <- post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_sd)

# individual item and SD variation
# all individual items/pt variation
vary_vals <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(vary_vals)
```

# section 2 #

## Visualise fixed effects ##

Using posterior_samples and tidybayes

```{r}
## fixed effects of interest, excluding basic intercepts for the ordinal model 
tidy_fixed <- post %>% 
  select(starts_with("b_")) %>% # select and rename in simpler labels
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(key, "targetd:attributed:groupd"), "target*attribute*group",
                if_else(str_detect(key, "attributed:groupd"), "attribute*group",
                if_else(str_detect(key, "targetd:groupd"), "target*group",
                if_else(str_detect(key, "targetd:attributed"), "target*attribute",
                if_else(str_detect(key, "targetd"), "target",
                if_else(str_detect(key, "attributed"), "attribute",
                if_else(str_detect(key, "groupd"), "group", "intercept"))))))),
         term = factor(term, levels=unique(term)))  
head(tidy_fixed)
tail(tidy_fixed)
str(tidy_fixed)

## data check
check.labels <- tidy_fixed %>% 
 group_by(key, term) %>% 
 distinct(key, term)
check.labels

## plot
p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term), 
                                       fill=term)) +  
  geom_vline(xintercept = 0, color = "red", alpha = 5/10) +
  stat_halfeye(alpha=0.7, point_size = 1) +
  labs(title = "Coefficient plot for fixed effects",
       x = NULL, y = NULL) +
  scale_fill_brewer(palette="Dark2") +
  theme(legend.position = "none")
p_tidy_fixed

ggsave ("exp2/figures/fixef.jpeg",
        width = 8, height = 6, dpi=800)
```

## make a table of fixed effects ##

```{r}
## make a df
fix_tab <- post_qi_b %>%
  mutate(key = factor(name, levels=unique(name)),
         term = if_else(str_detect(key, "targetd:attributed:groupd"), "target*attribute*group",
                if_else(str_detect(key, "attributed:groupd"), "attribute*group",
                if_else(str_detect(key, "targetd:groupd"), "target*group",
                if_else(str_detect(key, "targetd:attributed"), "target*attribute",
                if_else(str_detect(key, "targetd"), "target",
                if_else(str_detect(key, "attributed"), "attribute",
                if_else(str_detect(key, "groupd"), "group", "intercept"))))))),
         term = factor(term, levels=c("intercept", "target", "attribute", "group",
                                      "target*attribute", "target*group", "attribute*group",
                                      "target*attribute*group"))) %>% 
  arrange(term) %>%
  select(term, value, .lower, .upper) %>% 
  mutate(across(where(is.double), \(x) round(x, 2)))
fix_tab

## make a table using flextable (still needs a title but that can be fixed later, depending on the doc it is being used in)
p_fix_tab <- flextable(fix_tab) 
p_fix_tab

## save it
save_as_image(p_fix_tab, path = "exp2/tables/fixef.png")

## this table still needs polishing, but you get the basic idea
```

# section 3 #

## visualise varying effects ##

using posterior samples and tidybayes.

## variance at the group level (i.e., intercept for item and intercepts and effects across participants) ##

```{r}
## varying effects of interest, including item, as well as participant and effects across pts
tidy_vary <- post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(name = factor(name, levels=unique(name)),
         term = "intercept") 
head(tidy_vary)
tail(tidy_vary)

check.vary <- tidy_vary %>% 
  group_by(name, term) %>% 
  distinct(name, term)
check.vary

p_tidy_vary <- ggplot(tidy_vary, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "SD plot for varying effects",
       x = NULL, y = NULL) +
  scale_fill_brewer(palette="Dark2") +
  theme(legend.position = "none") 
p_tidy_vary

ggsave ("exp2/figures/vary.jpeg",
        width = 8, height = 6, dpi=800)
```

## variance within group levels i.e., at the individual item and participant level ##

Note - this should be in a centred metric, so that deviation from the group average (zero) can be visualised.

## filter and wrangle random effects within items and participants ##

## participant level ##

```{r}
# 99 pts x 1 term (intercept) = 99
vary_pt <- post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_pid"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(term = "intercept") %>% 
  arrange(term, Estimate) %>% 
  mutate(rank = rep(1:99, times=1),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
        outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pt)

check.vary.pt <- vary_pt %>% 
  distinct(param, term)
check.vary.pt
```

## plot varying effects via caterpillar plots ##

```{r}
p_tidy_varyvals <- ggplot(vary_pt, aes(x=rank, y=Estimate, colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying intercepts by participant",
       x = "participant")+
  theme(axis.text.x  = element_blank())+
  guides(colour = "none") +
  ylim(-0.5, 0.5)
p_tidy_varyvals

ggsave ("exp2/figures/varyvals.jpeg",
        width = 8, height = 6, dpi=800)
```

