---
title: "effects"
author: "Rich"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file visualises and tabulates parameters from Exp3 of the morality aesthetics project.

# load packages #

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork", "brms", 
         "tidybayes", "bayesplot", "future", "parallel", "flextable")

lapply(pkg, library, character.only = TRUE)
```

# adjust multicore settings #

```{r set-options}
options(mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()
```

# plot settings #

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face = "bold"), 
          title = element_text(size = 18, face = "bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# table settings #

```{r}
set_flextable_defaults(
  theme_fun = theme_vanilla,
  background.color = "white"
)
```

# read in previously saved brms models object files and/or fits #

If available, this saves time computing them each time 

```{r}
# bf <- readRDS("exp3/models/bf.rds") ## all data
bf_b1 <- readRDS("exp3/models/bf_b1.rds") ## block 1
# summary(bf_b1)
```

# read in the data #

```{r}
dataw_b1 <- read_csv("exp3/data/dataw_b1.csv") %>% 
  mutate(
    pid = factor(pid),
    item_no = factor(item_no), 
    condition = factor(condition, levels = c("factual", "low", "medium", "high"))
  ) %>%
  rename(
    item = item_no,
    status = art_status,
    gain = financial_gain
  )
head(dataw_b1)
str(dataw_b1)
```

# section 1 #

## Look at the point and interval estimates for the full model ##

using posterior_samples and tidybayes

```{r}
post <- as_draws_df(bf_b1) %>%  #
  select(-lp__) %>% 
  as_tibble()

post_summary <- posterior_summary(bf_b1, robust = TRUE)

# all fixed effects
post_qi_b <- post %>%
  select(starts_with("b_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_b)

# fixed effects (minus per item thresholds)
post_qi_fix <- post %>%
  select(starts_with("b_") & contains(c("condition"))) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_fix)

# varying effects - SD
post_qi_sd <- post %>%
  select(starts_with("sd_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(post_qi_sd)

# individual item and SD variation
# all individual items/pt variation
vary_vals <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  median_qi(value)
head(vary_vals)

# item variation
vary_vals_item <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  filter(str_detect(name, "item")) %>% 
  median_qi(value)
head(vary_vals_item)

# participant variation
vary_vals_pt <- post %>%
  select(starts_with("r_")) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  filter(str_detect(name, "pid")) %>% 
  median_qi(value)
head(vary_vals_pt)
```

# section 2 #

## Visualise fixed effects ##

Using posterior_samples and tidybayes

```{r}
## fixed effects of interest, excluding basic intercepts for the ordinal model 
tidy_fixed <- post %>% 
  select(starts_with("b_") & contains(c("condition"))) %>% # select and rename in simpler labels
  pivot_longer(everything()) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         dv = if_else(str_detect(key, "status"), "status",
              if_else(str_detect(key, "gain"), "gain", "prestige")),
         term = if_else(str_detect(key, "low"), "low",
                if_else(str_detect(key, "medium"), "medium", "high")),
         dv = factor(dv,
                     levels = c("gain", "prestige", "status")),
         term = factor(term, levels=unique(term)))  
head(tidy_fixed)
tail(tidy_fixed)
str(tidy_fixed)

## data check
check.labels <- tidy_fixed %>% 
 group_by(key, dv, term) %>% 
 distinct(key, dv, term)
check.labels

## plot
p_tidy_fixed <- ggplot(tidy_fixed, aes(x = value, y = fct_rev(term),
                                       fill=term)) +  
  geom_vline(xintercept = 0, color = "red", alpha = 5/10) +
  stat_halfeye(alpha = 0.7, point_size = 1) +
  labs(title = "Coefficient plot for fixed effects",
       x = NULL, y = NULL) +
  scale_fill_brewer(palette="Dark2") +
  theme(legend.position = "none")+
  facet_wrap(~dv, ncol = 3) +
  scale_x_continuous(limits = c(-3.5, 4), breaks=seq(-2, 4, 2))
p_tidy_fixed
```

## make a table of fixed effects ##

```{r}
## make a df
fix_tab <- post_qi_fix %>%
  mutate(key = factor(name, levels=unique(name)),
         dv = if_else(str_detect(key, "status"), "status",
              if_else(str_detect(key, "gain"), "gain", "prestige")),
         term = if_else(str_detect(key, "low"), "low",
                if_else(str_detect(key, "medium"), "medium", "high")),
         dv = factor(dv,
                     levels = c("gain", "prestige", "status")),
         term = factor(term, 
                       levels=c("low", "medium", "high")))   %>% 
  arrange(dv, term) %>%
  select(dv, term, value, .lower, .upper) %>% 
  mutate(across(where(is.double), \(x) round(x, 2)))
fix_tab

## make a table using flextable (still needs a title but that can be fixed later, depending on the doc it is being used in)
p_fix_tab <- flextable(fix_tab) 
p_fix_tab

## save it
save_as_image(p_fix_tab, path = "exp3/tables/fixef.png")

## this table still needs polishing, but you get the basic idea
```

## compare levels of condition ##

The above fixed effects are estimates that compare low, medium and high to the factual condition. 
That is, the factual condition is the reference category.
Here we might compare levels between low, medium and high also.

use compare_levels from tidybayes

now use tidy_draws

```{r}
## it is not necessary, but to make it easier with compare levels, first take draws with tidybayes
## really, we could use this above instead of as_draws_df()
tidy_post <- tidy_draws(bf_b1) %>% 
  select(.chain, .iteration, .draw,
         starts_with("b_") & contains(c("condition"))) %>% # 
  pivot_longer(-c(".chain", ".iteration", ".draw")) %>% # move from wide to long
  mutate(key = factor(name, levels=unique(name)),
         dv = if_else(str_detect(key, "status"), "status",
              if_else(str_detect(key, "gain"), "gain", "prestige")),
         term = if_else(str_detect(key, "low"), "low",
                if_else(str_detect(key, "medium"), "medium", "high")),
         dv = factor(dv,
                     levels = c("gain", "prestige", "status")),
         term = factor(term = factor(term, 
                       levels=c("low", "medium", "high")))) 
head(tidy_post)
tail(tidy_post)
str(tidy_post)

check.labels <- tidy_post %>% 
 group_by(key, dv, term) %>% 
 distinct(key, dv, term)
check.labels 
```

what about compare levels?

this might take a few minutes...

```{r}
plan(multicore)
tidy_diffs <- tidy_post %>%
  select(-key, -name) %>% 
  group_by(.chain, .iteration, .draw, dv) %>% 
  compare_levels(variable = value,
                 by = term)
tidy_diffs
```

save out tidy diffs

```{r}
write_csv(tidy_diffs, "exp3/draws/tidy_diffs.csv")
```

plot

```{r}
p_tidy_diff <- ggplot(tidy_diffs, aes(x = value, y = fct_rev(term))) +
  geom_vline(xintercept = 0, color = "red", alpha = 5/10) +
  stat_halfeye(alpha = 0.7, point_size = 1) +
  labs(title = "Comparing levels of condition",
       x = NULL, y = NULL) +
  scale_fill_manual(values = "darkgrey") +
  theme(legend.position = "none")+
  facet_wrap(~dv, ncol = 3) +
  scale_x_continuous(limits = c(-3.5, 4), breaks=seq(-2, 4, 2))
p_tidy_diff
```

summarise with quantiles

```{r}
tidy_diffs_q <- tidy_diffs %>% 
  group_by(dv, term) %>% 
  median_qi(value)
tidy_diffs_q
```

make a table of comparisons

```{r}
diff_tab <- tidy_diffs_q %>% 
  select(dv, term, value, .lower, .upper) %>% 
  mutate(across(where(is.double), \(x) round(x, 2)))
diff_tab

p_diff_tab <- flextable(diff_tab) 
p_diff_tab

## save it
save_as_image(p_diff_tab, path = "exp3/tables/diffs.png")
```

plot the fixed effects and level comparisons together

```{r}
p2.1 <- (p_tidy_fixed / p_tidy_diff) +
  plot_annotation(tag_levels = 'A')
p2.1

ggsave ("exp3/figures/fixef_diff_b1.jpeg",
        width = 10, height = 7, dpi=800)
```


# section 3 #

## visualise varying effects ##

using posterior samples and tidybayes.

## variance at the group level (i.e., intercept for item and intercepts and effects across participants) ##

```{r}
## varying effects of interest, including item, as well as participant and effects across pts
tidy_vary <- post %>% 
  select(contains("sd_")) %>% 
  pivot_longer(everything()) %>% 
  mutate(key = factor(name, levels=unique(name)),
         group = if_else(str_detect(key, "item"), "item", "pid"),
         term = if_else(str_detect(key, "Intercept") 
                        & group == "item","item_intercept",
                if_else(str_detect(key, "Intercept") 
                        & group == "pid","pid_intercept",
                if_else(str_detect(key, "low"), "low",
                if_else(str_detect(key, "medium"), "medium", "high")))),
         dv = if_else(str_detect(key, "status"), "status",
              if_else(str_detect(key, "gain"), "gain", "prestige")),
         group = factor(group, levels=unique(group)),
         term = factor(term, levels=unique(term)),
         dv = factor(dv,levels=unique(dv))) %>% 
  arrange(group)
head(tidy_vary)
tail(tidy_vary)

check.vary <- tidy_vary %>% 
  group_by(key, group, term, dv) %>% 
  distinct(key, group, term, dv)
check.vary

p_tidy_vary <- ggplot(tidy_vary, aes(x = value, y = fct_rev(term), fill=term)) +  
  geom_vline(xintercept = 0, color = "grey", alpha = 5/10) +
  stat_halfeye() +
  labs(title = "SD plot for varying effects",
       x = NULL, y = NULL) +
  scale_fill_brewer(palette="Dark2") +
  theme(panel.grid   = element_blank(),
        panel.grid.major.y = element_line(color = alpha("firebrick4", 1/2), 
                                          linetype = 3),
        # axis.text.y  = element_blank(),
        # axis.ticks.y = element_blank(),
        legend.position = "none") +
  # coord_cartesian(xlim =c(0, 1.5)) +
  # scale_x_continuous(breaks=seq(0,1.5,0.5)) +
  facet_wrap(~dv,ncol=3)
p_tidy_vary

ggsave ("exp3/figures/vary_b1.jpeg",
        width = 8, height = 6)
```

## variance within group levels i.e., at the individual item and participant level ##

Note - this should be in a centred metric, so that deviation from the group average (zero) can be visualised.

## filter and wrangle random effects within items and participants ##

## item level ##

```{r}
# item level
# using post_summary (from posterior_summary)
# 40 items x 3 dvs = 120
vary_item <- post_summary %>% 
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_item"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(dv = if_else(str_detect(param, "status"), "status",
              if_else(str_detect(param, "gain"), "gain", "prestige")),
         term = "item_intercept",
         dv = factor(dv, levels = unique(dv))) %>% 
  arrange(dv, Estimate) %>% 
  mutate(rank = rep(1:40, times=3),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
         outlier = factor(outlier, levels = c("no", "yes")))
head(vary_item)
```

## participant level ##

```{r}
# 100 pts x 4 terms x 3 dvs = 1200
vary_pt <- post_summary %>%
  data.frame() %>% 
  rownames_to_column("param") %>% 
  filter(str_detect(param, "r_pid"),
         str_detect(param, "cor", negate = TRUE)) %>% 
  mutate(dv = if_else(str_detect(param, "status"), "status",
              if_else(str_detect(param, "gain"), "gain", "prestige")),
         term = if_else(str_detect(param, "Intercept"), "pid_intercept",
                if_else(str_detect(param, "low"), "low",
                if_else(str_detect(param, "medium"), "medium", "high"))),
         dv = factor(dv, levels = unique(dv)),
         term = factor(term, levels = unique(term))) %>% 
  arrange(dv, term, Estimate) %>% 
  mutate(rank = rep(1:100, times=12),
         outlier = if_else(Q2.5 > 0 | Q97.5 < 0, "yes", "no"), 
        outlier = factor(outlier, levels = c("no", "yes")))
head(vary_pt)

check.vary.pt <- vary_pt %>% 
  distinct(dv, term)
check.vary.pt
```

## join item and pt varyvals together ##

```{r}
tidy_varyvals <- vary_item %>% 
  union(vary_pt) %>% 
  mutate(term = factor(term, levels = unique(term)))
head(tidy_varyvals)
```

## plot varying effects via caterpillar plots ##

```{r}
p_tidy_varyvals <- ggplot(tidy_varyvals, aes(x=rank, y=Estimate, colour=outlier))+
  geom_pointrange(aes(ymin=Q2.5,ymax=Q97.5),fatten = 0.5)+
  geom_hline(yintercept=0, colour = "red")+
  theme_bw()+
  scale_colour_manual(values=c("darkgrey","black")) +
  labs(title = "Varying effects by stimulus item and participant",
       x = "stimulus item / participant")+
  theme(axis.text.x  = element_blank())+
  guides(colour = "none")+
  facet_grid(dv~term, scales = "free_x")
p_tidy_varyvals

ggsave ("exp3/figures/varyvals_b1.jpeg",
        width = 8, height = 6)
```

