---
title: "model"
author: "Rich & Ionela"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file builds some models for Exp3 in the morality aesthetics project.

# load the libraries that we will be using #

## load ##

```{r load-pkg}
pkg <- c("cmdstanr", "standist", "tidyverse", "RColorBrewer", "patchwork", 
         "brms", "tidybayes", "bayesplot", "future", "parallel")

lapply(pkg, library, character.only = TRUE)
```

## settings ##

```{r set-options}
options(brms.backend = "cmdstanr",
        mc.cores = parallel::detectCores(),
        future.fork.enable = TRUE,
        future.rng.onMisuse = "ignore") ## automatically set in RStudio

supportsMulticore()

detectCores()
```

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face="bold"), 
          title = element_text(size = 18, face="bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# section 1 #

## read in data and create factors where necessary ##

all the data

```{r}
dataw <- read_csv("exp3/data/dataw.csv") %>% 
  mutate(
    pid = factor(pid),
    item_no = factor(item_no), 
    condition = factor(condition, levels = c("factual", "low", "medium", "high"))
  ) %>%
  rename(
    item = item_no,
    status = art_status,
    gain = financial_gain
  )
head(dataw)
str(dataw)
```

just block 1 data

```{r}
dataw_b1 <- read_csv("exp3/data/dataw_b1.csv") %>% 
  mutate(
    pid = factor(pid),
    item_no = factor(item_no), 
    condition = factor(condition, levels = c("factual", "low", "medium", "high"))
  ) %>%
  rename(
    item = item_no,
    status = art_status,
    gain = financial_gain
  )
head(dataw_b1)
str(dataw_b1)
```

# section 2 #

## build some models ##

## b0.1 - intercepts only ##

## formula ##

```{r}
formula = bf(mvbind(status, gain, prestige) ~ 1) + set_rescor(FALSE)
```

## check the priors available ##

```{r}
get_prior(formula,
          data = dataw, family = cumulative("probit"))
```

## set priors ##

```{r}
# set priors #
priors <- set_prior("normal(0, 1)", class = "Intercept",
                    resp = c("status", "gain", "prestige"))
```

## run the model ##

```{r}
plan(multicore)
b0.1 <- brm(formula = formula,
        data = dataw, family = cumulative("probit"),
        prior = priors,
        iter = 5000, warmup = 1000, cores = 20, chains = 4,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        save_pars = save_pars(all=TRUE),
        seed = 123,
        init = 0.1,
        file = "exp3/models/b0.1")
summary(b0.1)
```

## bf - full model ##

## formula ##

```{r}
formula <- bf(mvbind(status, gain, prestige) | thres(4, gr = item) ~ 
                1 + condition +
                (1 | p | item) +
                (1 + condition | a | pid)) + 
                set_rescor(FALSE)
```

## check the priors available ##

```{r}
get_prior(formula,
          data = dataw, family = cumulative("probit"))
```

## set priors ##

```{r}
# set priors #
priors <- c(
  set_prior('normal(0, 1.5)', class = 'Intercept',
            resp = c("status", "gain", "prestige")),
    set_prior('normal(0, 1)', class = 'b',
            resp = c("status", "gain", "prestige")),
    set_prior('normal(0, 1)', class = 'sd',
            resp = c("status", "gain", "prestige")),
    set_prior('lkj(2)', class = 'cor')
)
```

## run the model ##

```{r}
# this will help us track time
t1 <- Sys.time()

plan(multicore)
bf <- brm(formula = formula,
        data = dataw, family = cumulative("probit"),
        prior = priors,
        iter = 5000, warmup = 1000, cores = 20, chains = 4,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        save_pars = save_pars(all=TRUE),
        seed = 123,
        init = 0.1,
        file = "exp3/models/bf")
summary(bf)

t2 <- Sys.time()

t2 - t1

# Time difference of ~5 hours imac.
```

take a look

pp_check 

```{r}
# (just the understanding dv for now to take a look)
pp_bf <- pp_check(bf, ndraws = 100, resp = "use")
pp_bf
```

chains

```{r}
plot(bf)
```


## bf_b1 - full model with block 1 data only ##

## formula ##

```{r}
formula <- bf(mvbind(status, gain, prestige) | thres(4, gr = item) ~ 
                1 + condition +
                (1 | p | item) +
                (1 + condition | a | pid)) + 
                set_rescor(FALSE)
```

## check the priors available ##

```{r}
get_prior(formula,
          data = dataw_b1, family = cumulative("probit"))
```

## set priors ##

```{r}
# set priors #
priors <- c(
  set_prior('normal(0, 1.5)', class = 'Intercept',
            resp = c("status", "gain", "prestige")),
    set_prior('normal(0, 1)', class = 'b',
            resp = c("status", "gain", "prestige")),
    set_prior('normal(0, 1)', class = 'sd',
            resp = c("status", "gain", "prestige")),
    set_prior('lkj(2)', class = 'cor')
)
```

## run the model ##

```{r}
# this will help us track time
t1 <- Sys.time()

plan(multicore)
bf_b1 <- brm(formula = formula,
        data = dataw_b1, family = cumulative("probit"),
        prior = priors,
        iter = 5000, warmup = 1000, cores = 20, chains = 4,
        control = list(adapt_delta = 0.99, max_treedepth = 15),
        save_pars = save_pars(all=TRUE),
        seed = 123,
        init = 0.1,
        file = "exp3/models/bf_b1")
summary(bf_b1)

t2 <- Sys.time()

t2 - t1

# Time difference of ~2 hours imac.
```

take a look

pp_check 

```{r}
# (just the understanding dv for now to take a look)
pp_bf_b1 <- pp_check(bf_b1, ndraws = 100, resp = "use")
pp_bf_b1
```

chains

```{r}
plot(bf_b1)
```