---
title: "wrangle"
author: "Rich & Ionela"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file wrangles raw data from the morality AI-generated art project (Experiment 3), which involved rating data (on likert scales). It produces some summary data plots, saves out data files for modelling and further analysis in later scripts.

# load the libraries that we will be using #

## install ##

```{r install-pkg}
# install.packages(c("tidyverse", "RColorBrewer", "patchwork"))
```

take a snapshot of loaded packages and update the lock.file using renv

```{r snapshot-renv}
# take a snapshot and update the lock.file
# renv::snapshot() # this is only necessary when new packages or installed or packages are updated.
```

## load ##

```{r load-pkg}
pkg <- c("tidyverse", "RColorBrewer", "patchwork")

lapply(pkg, library, character.only = TRUE)
```

## plot settings ##

theme settings for ggplot

```{r, eval = F}
theme_set(
  theme_bw() +
    theme(text = element_text(size = 18, face="bold"), 
          title = element_text(size = 18, face="bold"),
          legend.position = "bottom")
)

## Set the amount of dodge in figures
pd <- position_dodge(0.7)
pd2 <- position_dodge(1)
```

# section 1 #

## read in previously wrangled data, if necessary ##

this is good for plotting, if you've already wrangled and saved the data before.

If you haven't already wrangled and saved, then skip this chunk.

```{r}

```

## read in data and select the variables we need ##

```{r}
raw_data <- read_csv("exp3/data/moral_exp3.csv") %>%
  rename(pid = Participant) %>% # rename
  select(pid,fact1_ArtStatus:high10_Prestige,fact_img1:high_img10,failedCheck)
head(raw_data)
```

# section 2 #

## do some wrangling ##

### filter by attention checks. According to our pre-registration we should remove ppts who have failed 3 attention checks ###

```{r}
raw_data %>% 
  ggplot(aes(x="", y=failedCheck)) +
  geom_jitter(height = 0)
```

Everyone passed the required attention checks

```{r}
att_checks <- raw_data %>%
  filter(failedCheck == 3) %>%
  select(pid, failedCheck)
head(att_checks)
```

### re-shape the data into different formats ###

```{r}
data_long <- raw_data %>%
  select(pid, matches("ArtStatus$|FinanGain$|Prestige$")) %>%  
  pivot_longer(
    cols = -pid,  
    names_to = c("condition", "dv"),  
    names_sep = "_",  
    values_to = "ratings"  
  ) %>%
  mutate(
    # Remove numeric from condition (e.g., fact1 -> fact, low1 -> low, etc.)
    condition = str_replace(condition, "\\d+$", ""),  # Remove digits at the end of the string
    # raname fact as factual
    condition = if_else(condition == "fact", "factual", condition),
    # Recoding dv 
    dv = recode(dv, FinanGain = "financial_gain", Prestige = "prestige", ArtStatus = "art_status"),
    condition = factor(condition, levels = c("factual", "low", "medium", "high")),
    dv = factor(dv, levels = c("financial_gain", "prestige", "art_status"))
  )
head(data_long)
str(data_long)
summary(data_long)
```

### re-shape the item data into long format ###

```{r}
data_long_item <- raw_data %>%
  select(pid, matches("img")) %>%  
  pivot_longer(
    cols = -pid,  
    names_to = c("condition", "item"),  
    names_sep = "_",  
    values_to = "item_no"  
  ) %>% 
  mutate(condition = if_else(condition == "fact", "factual", condition))
head(data_long_item)
```

### re-shape the long morality data and item long data into one long format ###

```{r}
# take item_no from item_data_long and replicate each item_no 3 times for each of the 3 dvs
item_sequence <- rep(data_long_item$item_no, each = 3)

# add the item_no_sequence as a new column to data_long without changing row order
datal <- data_long %>%
  mutate(item_no = item_sequence)
head(datal)
```

data check

```{r}
data.check0 <- datal %>% 
  group_by(condition, dv) %>%
  tally()
head(data.check0)
```

convert from long into wide format (this is useful for modelling each dv separately)

```{r}
dataw <- datal %>% 
  pivot_wider(names_from = "dv", 
              values_from = "ratings")
head(dataw)
str(dataw)
summary(dataw)
```

# section 3 #

## calculate summary statistics ##

```{r}
summary_pid <- datal %>%
   group_by(pid, condition, dv) %>%
   summarise(mean=mean(ratings, na.rm = TRUE),
             n=n(),
             sd=sd(ratings, na.rm = TRUE),
             sem=sd/sqrt(n))  %>%
   ungroup() 
head(summary_pid)
```

summary data at the group level

```{r}
# by condition, dv
summary <- datal %>% 
  group_by(condition, dv) %>%
  summarise(mean=mean(ratings, na.rm = TRUE),
            n=length(unique(pid)),
            sd=sd(ratings, na.rm = TRUE),
            sem=sd/sqrt(n))  %>%
   ungroup() 
summary
```

# section 4 #

## plot ##

make some facet labels without underscores (this makes Emily happy...)

```{r}
# facet labels for dv
dv_labels <- c(
  "financial_gain" = "Financial Gain",
  "prestige" = "Prestige",
  "art_status" = "Art Status"
)
```

violin

```{r}
p4.1 <- ggplot(summary, 
              aes(x = condition, y = mean, fill = condition)) + 
   geom_jitter(data = summary_pid, position = position_jitterdodge(dodge.width = 1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(data = summary_pid, alpha = 0.7, position = pd2) +
   geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .2, position = pd2) +
   geom_line(aes(group = dv), position = pd2) +
   geom_point(position = pd2, size = 3) +
   scale_colour_brewer(palette = "Accent") +
   scale_fill_brewer(palette = "Accent") +
   labs(x = "", y = "Ratings (1-5)") +
   facet_grid(~ dv, labeller = labeller(dv = dv_labels)) +
   ggtitle("Ratings by Condition")
p4.1

# Save the plot
ggsave("exp3/figures/ratings_violin.jpeg", 
       width = 10, height = 6, dpi = 800)
```

try a different plot - flip the axes and use point and error bars

```{r}
p4.2 <- ggplot(summary, 
              aes(y = condition, x = mean, fill = condition, colour = condition)) + 
   geom_point(position = position_dodge(width = 0.7), size = 3) +  
   geom_errorbar(aes(xmin = mean - sem, xmax = mean + sem, colour = condition), 
                 width = 0.2, position = position_dodge(width = 0.7), linewidth = 1) +  
   geom_vline(data=filter(summary, condition=="factual"), aes(xintercept=mean), colour="black",
              linetype = "dashed") +
   scale_x_continuous(limits = c(1, 3.5), breaks = seq(1, 3, by = 1)) +  # Set x-axis from 1 to 5
   scale_colour_brewer(palette = "Set2") +  
   scale_fill_brewer(palette = "Set2") +
   labs(x = "Ratings (1-5)", y = "") +
   facet_wrap(~dv, labeller = labeller(dv = dv_labels)) +
   theme(legend.position = "none") +  
   ggtitle("Ratings by Condition")
p4.2

ggsave("exp3/figures/ratings_point.jpeg",
       width = 10, height = 6, dpi = 800)
```

density

```{r}
p4.3 <- ggplot(datal, aes(x = ratings, fill = condition)) +
  geom_density(alpha = 0.7) +
  scale_colour_brewer(palette = "Accent") +
  scale_fill_brewer(palette = "Accent") +
  facet_grid(condition ~ dv, labeller = labeller(dv = dv_labels)) +  # Use custom labels for dv
  theme(
    legend.position = "none"                
  ) +
  labs(x = "Ratings", y = "Density")
p4.3

# ggsave("figures/ratings_density.jpeg", 
#        width = 10, height = 6, dpi = 800)
```

now overlay condition?

```{r}
p4.4 <- ggplot(datal, aes(x = ratings, fill = condition,
                          colour = condition)) +
  geom_density(alpha = 0.5) +
  scale_colour_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(~dv, labeller = labeller(dv = dv_labels)) +  # Use custom labels for dv
  labs(x = "Ratings", y = "Density")
p4.4
```

# section 5 #

## look at order effects ##

the design involved a partially fixed order of conditions, such that the factual condition was always first, then the remaining conditions were randomised/counterbalanced.

Here, we take a look at possible order effects, rather than condition effects. 

```{r}
# raw_order <- read_csv("exp3/data/moral_exp3_block_order.csv") %>%
#   rename(pid = Participant) %>% # rename
#   select(pid,fact1_ArtStatus:high10_Prestige,block_order_1:block_order_3)
# head(raw_order)

order <- read_csv("exp3/data/moral_exp3_block_order.csv") %>%
  rename(pid = Participant) %>% # rename
  select(pid,,block_order_1:block_order_3) %>% 
  add_column(block_order_0 = "factual") %>% 
  relocate(block_order_0, .after = pid) %>% 
  mutate(sequence = case_when(
    block_order_1 == "low" &  block_order_2 == "medium" & block_order_3 == "high" ~ "lmh",
    block_order_1 == "low" &  block_order_2 == "high" & block_order_3 == "medium" ~ "lhm",
    block_order_1 == "medium" &  block_order_2 == "low" & block_order_3 == "high" ~ "mlh",
    block_order_1 == "medium" &  block_order_2 == "high" & block_order_3 == "low" ~ "mhl",
    block_order_1 == "high" &  block_order_2 == "low" & block_order_3 == "medium" ~ "hlm",
    .default = "hml"
  )) %>% 
  pivot_longer(cols = block_order_0: block_order_3,
               names_to = "block_order",
               values_to = "condition")
head(order)
```

join order with ratings data

```{r}
datal_order <- datal %>% 
  inner_join(order, by = c("pid", "condition")) %>% 
  mutate(block_order = as.numeric(str_remove(block_order, "block_order_")),
         block_order = factor(block_order),
         sequence = factor(sequence,
                           levels = c("lmh", "lhm", "mlh", "mhl", "hlm", "hml")))
head(datal_order)
str(datal_order)
```

## take a look at block order effects (on average across sequences) ##

summary stats by block order (i.e., 0,1,2,3)

```{r}
summary_order_pid <- datal_order %>%
   group_by(pid, block_order, dv) %>%
   summarise(mean=mean(ratings, na.rm = TRUE),
             n=n(),
             sd=sd(ratings, na.rm = TRUE),
             sem=sd/sqrt(n))  %>%
   ungroup() 
head(summary_order_pid)
```

summary data at the group level

```{r}
# by condition, dv
summary_order <- datal_order %>% 
  group_by(block_order, dv) %>%
  summarise(mean=mean(ratings, na.rm = TRUE),
            n=length(unique(pid)),
            sd=sd(ratings, na.rm = TRUE),
            sem=sd/sqrt(n))  %>%
   ungroup() 
summary_order
```

plot

violin

```{r}
p5.1 <- ggplot(summary_order, 
              aes(x = block_order, y = mean, fill = block_order)) + 
   geom_jitter(data = summary_order_pid, position = position_jitterdodge(dodge.width = 1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(data = summary_order_pid, alpha = 0.7, position = pd2) +
   geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .2, position = pd2) +
   geom_line(aes(group = dv), position = pd2) +
   geom_point(position = pd2, size = 3) +
   scale_colour_brewer(palette = "Accent") +
   scale_fill_brewer(palette = "Accent") +
   labs(x = "", y = "Ratings (1-5)") +
   facet_wrap(~dv, labeller = labeller(dv = dv_labels)) +
   ggtitle("Ratings by Block Order")
p5.1

# Save the plot
ggsave("exp3/figures/ratings_order_violin.jpeg", 
       width = 10, height = 6, dpi = 800)
```

try a different plot - flip the axes and use point and error bars

```{r}
p5.2 <- ggplot(summary_order, 
              aes(y = block_order, x = mean, fill = block_order, colour = block_order)) + 
   geom_point(position = position_dodge(width = 0.7), size = 3) +  
   geom_errorbar(aes(xmin = mean - sem, xmax = mean + sem, colour = block_order), 
                 width = 0.2, position = position_dodge(width = 0.7), linewidth = 1) +  
   geom_vline(data=filter(summary_order, block_order=="0"), aes(xintercept=mean), colour="black",
              linetype = "dashed") +
   scale_x_continuous(limits = c(1, 3.5), breaks = seq(1, 3, by = 1)) +  # Set x-axis from 1 to 5
   scale_colour_brewer(palette = "Set2") +  
   scale_fill_brewer(palette = "Set2") +
   labs(x = "Ratings (1-5)", y = "") +
   facet_wrap(~dv, labeller = labeller(dv = dv_labels)) +
   theme(legend.position = "none") +  
   ggtitle("Ratings by Block Order")
p5.2

ggsave("exp3/figures/ratings_order_point.jpeg",
       width = 10, height = 6, dpi = 800)
```

## take a look at blok order effects separated by sequece type ##

summary stats by block order (i.e., 0,1,2,3) and block sequence 

at the pid level

```{r}
summary_seq_pid <- datal_order %>%
   group_by(pid, block_order, sequence, dv) %>%
   summarise(mean=mean(ratings, na.rm = TRUE),
             n=n(),
             sd=sd(ratings, na.rm = TRUE),
             sem=sd/sqrt(n))  %>%
   ungroup() 
head(summary_seq_pid)
```

summary data at the group level

```{r}
# by condition, dv
summary_seq <- datal_order %>% 
  group_by(block_order, sequence, dv) %>%
  summarise(mean=mean(ratings, na.rm = TRUE),
            n=length(unique(pid)),
            sd=sd(ratings, na.rm = TRUE),
            sem=sd/sqrt(n))  %>%
   ungroup() 
summary_seq
```

plot

violin

```{r}
p5.3 <- ggplot(summary_seq, 
              aes(x = block_order, y = mean, fill = block_order)) + 
   geom_jitter(data = summary_seq_pid, position = position_jitterdodge(dodge.width = 1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(data = summary_seq_pid, alpha = 0.7, position = pd2) +
   geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .2, position = pd2) +
   geom_line(aes(group = dv), position = pd2) +
   geom_point(position = pd2, size = 3) +
   scale_colour_brewer(palette = "Accent") +
   scale_fill_brewer(palette = "Accent") +
   labs(x = "", y = "Ratings (1-5)") +
   facet_grid(sequence~dv, labeller = labeller(dv = dv_labels)) +
   ggtitle("Ratings by Block Order and Block Sequence")
p5.3

# Save the plot
# ggsave("exp3/figures/ratings_seq_violin.jpeg", 
#        width = 10, height = 6, dpi = 800)
```

try a different plot - flip the axes and use point and error bars

```{r}
p5.4 <- ggplot(summary_seq, 
              aes(y = block_order, x = mean, fill = block_order, colour = block_order)) + 
   geom_point(position = position_dodge(width = 0.7), size = 3) +  
   geom_errorbar(aes(xmin = mean - sem, xmax = mean + sem, colour = block_order), 
                 width = 0.2, position = position_dodge(width = 0.7), linewidth = 1) +  
   geom_vline(data=filter(summary_seq, block_order=="0"), aes(xintercept=mean), colour="black",
              linetype = "dashed") +
   scale_x_continuous(limits = c(1, 3.5), breaks = seq(1, 3, by = 1)) +  # Set x-axis from 1 to 5
   scale_colour_brewer(palette = "Set2") +  
   scale_fill_brewer(palette = "Set2") +
   labs(x = "Ratings (1-5)", y = "") +
   facet_grid(sequence~dv, labeller = labeller(dv = dv_labels)) +
   theme(legend.position = "none") +  
   ggtitle("Ratings by Block Order and Block Sequence")
p5.4

ggsave("exp3/figures/ratings_seq_point.jpeg",
       width = 10, height = 8, dpi = 800)
```

# section 6 #

## only look at the baseline and block 1 ##

The aim here is to look at a subsample of the data that is not affected by order effects.
So, here, we look at the baseline (factual) condition and then compare to all other conditions from block 1.
That way, we effectively turn the experiment into a within and between design.
There are within-group comparisons between conditions (factual vs low, factual vs medium and factual vs high).
Then there could be between group comparisons (low vs medium, low vs high etc.)
This makes sure that all participants at this point in the experiment have only been exposed to two conditions (factual and one of the following: low, medium and high).

## do some wrangling ##

```{r}
datal_order_b1 <- datal_order %>% 
  filter(block_order %in% c(0, 1)) %>% 
  mutate(condition = factor(condition,
                            levels = c("factual", "low", "medium", "high")))
datal_order_b1
```

## summary stats ##

summary stats by block order (i.e., 0,1)

at the pid level

```{r}
summary_pid_b1 <- datal_order_b1 %>%
   group_by(pid, block_order, condition, dv) %>%
   summarise(mean=mean(ratings, na.rm = TRUE),
             n=n(),
             sd=sd(ratings, na.rm = TRUE),
             sem=sd/sqrt(n))  %>%
   ungroup() 
head(summary_pid_b1)
```

summary data at the group level

```{r}
# by condition, dv
summary_b1 <- datal_order_b1 %>% 
  group_by(block_order, condition, dv) %>%
  summarise(mean=mean(ratings, na.rm = TRUE),
            n=length(unique(pid)),
            sd=sd(ratings, na.rm = TRUE),
            sem=sd/sqrt(n))  %>%
   ungroup() 
summary_b1
```

## plot by condition ##

violin

```{r}
p6.1 <- ggplot(summary_b1, 
              aes(x = condition, y = mean, fill = condition)) + 
   geom_jitter(data = summary_pid_b1, position = position_jitterdodge(dodge.width = 1), 
               alpha = 1, colour = "darkgrey") +
   geom_violin(data = summary_pid_b1, alpha = 0.7, position = pd2) +
   geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .2, position = pd2) +
   geom_line(aes(group = dv), position = pd2) +
   geom_point(position = pd2, size = 3) +
   scale_colour_brewer(palette = "Accent") +
   scale_fill_brewer(palette = "Accent") +
   labs(x = "", y = "Ratings (1-5)") +
   facet_grid(~ dv, labeller = labeller(dv = dv_labels)) +
   ggtitle("Ratings by Condition (Block 1)")
p6.1

# Save the plot
ggsave("exp3/figures/ratings_violin_b1.jpeg", 
       width = 10, height = 6, dpi = 800)
```

try a different plot - flip the axes and use point and error bars

```{r}
p6.2 <- ggplot(summary_b1, 
              aes(y = condition, x = mean, fill = condition, colour = condition)) + 
   geom_point(position = position_dodge(width = 0.7), size = 3) +  
   geom_errorbar(aes(xmin = mean - sem, xmax = mean + sem, colour = condition), 
                 width = 0.2, position = position_dodge(width = 0.7), linewidth = 1) +  
   geom_vline(data=filter(summary_b1, condition=="factual"), aes(xintercept=mean), colour="black",
              linetype = "dashed") +
   scale_x_continuous(limits = c(1, 3.5), breaks = seq(1, 3, by = 1)) +  # Set x-axis from 1 to 5
   scale_colour_brewer(palette = "Set2") +  
   scale_fill_brewer(palette = "Set2") +
   labs(x = "Ratings (1-5)", y = "") +
   facet_wrap(~dv, labeller = labeller(dv = dv_labels)) +
   theme(legend.position = "none") +  
   ggtitle("Ratings by Condition (Block 1)")
p6.2

ggsave("exp3/figures/ratings_point_b1.jpeg",
       width = 10, height = 6, dpi = 800)
```

## create a wide dataset for modelling ##

```{r}
dataw_b1 <- datal_order_b1 %>% 
  select(pid, condition, item_no, dv, ratings) %>% 
  pivot_wider(id_cols = c(pid, condition, item_no),
              names_from = "dv",
              values_from = "ratings")
head(dataw_b1)
```


# section 7 #

## save data files ##

```{r}
# data in long and wide formats
write_csv(datal, "exp3/data/datal.csv")
write_csv(dataw, "exp3/data/dataw.csv")
write_csv(datal_order, "exp3/data/datal_order.csv")
write_csv(dataw_b1, "exp3/data/dataw_b1.csv")

# summary statistics at the group level in long format
write_csv(summary, "exp3/data/summary.csv")
write_csv(summary_order, "exp3/data/summary_order.csv")
write_csv(summary_seq, "exp3/data/summary_seq.csv")

# and just the order information per pid
write_csv(order, "exp3/data/order.csv")
```